---
title: "Case Study 1"
author: "Chase Darlington"
date: "October 3, 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#for MAC
load("~/Downloads/ramen.Rdata")
```


```{r}
#for PC
#load("C:\\Users\\Chase Darlington\\Downloads\\ramen.RData")

#ramen
#names(ramen)
```

1) (3 pts) How many different brands are reviewed in the data set? Do you recognize any of them?

355 brands are reviewed in the data set.
```{r, }
#rawbrands <- ramen$Brand
#brands <- data.frame(rawbrands[!duplicated(rawbrands)])
#brands
#nrow(brands)
#nrow(ramen$Brand[!duplicated(ramen$Brand)])

#or

length(ramen$Brand[!duplicated(ramen$Brand)])

```


2) (3 pts) What years do we have "top ten" data from? (Hint: This will require turning the Top Ten
column into a string column and using some of our string functions)

We have top ten information from 2012-2016
```{r}
ramen$Top.Ten[ramen$Top.Ten==""] <- NA
Top.Ten <- na.omit(ramen)
#Top.Ten

seperatedTop.Ten <- data.frame(do.call('rbind', strsplit(as.character(Top.Ten$Top.Ten), " #", fixed=TRUE)))
#seperatedTop.Ten
levels(seperatedTop.Ten$X1)
```


3) (3 pts) Which ramen brands are from the United States (hint: beware of abbreviations!)?

Refer below for the 49 brands available in the United States
```{r}
#levels(ramen$Country)
sort(ramen$Country[!duplicated(ramen$Country)])
#Note: there exists a USA and United States
#So, reassign values
ramen$Country[ramen$Country=="United States"] <- "USA"
#Check 3. For some reason USA is still listed as a level...
#levels(ramen$Country)
#...so, Check 2
ramen[ramen$Country=="United States"]
ramen$Country[ramen$Country=="United States"]
#reprint all the countries; no duplicates
#sort(ramen$Country[!duplicated(ramen$Country)])
#levels(ramen$Country)



#Then, Dubai isn't a country, nor Sarawak: correct to UAE, and Malaysia; Hong Kong is a special administrative region, so it remains.
ramen$Country[ramen$Country=="Dubai"] <- "UAE"
ramen$Country[ramen$Country=="Sarawak"] <- "Malaysia"
#sort(ramen$Country[!duplicated(ramen$Country)])
```

```{r}
#notably, all United States entries were converted to USA
USbrands <- (ramen[ramen$Country=="USA",])
USbrands <- (USbrands[!duplicated(USbrands$Brand),])
length(USbrands$Brand)
data.frame(USbrands$Brand)
data.frame(sort(USbrands$Brand))

```

4) (3 pts) Has any brand of ramen won the #1 slot in the Top Ten Ramens list more than once? Which
brand(s)?

MyKuali ranked #1 2014 to 2015, 2 years in a row

```{r}
#reapply NA,just in case
ramen$Top.Ten[ramen$Top.Ten==""] <- NA
Top.Ten <- na.omit(ramen)
#Top.Ten

seperatedTop.Ten <- data.frame(do.call('rbind', strsplit(as.character(Top.Ten$Top.Ten), " #", fixed=TRUE)))
class(seperatedTop.Ten)
colnames(seperatedTop.Ten) <- c("Year","Rank")
#seperatedTop.Ten
Top.Ten <- cbind(Top.Ten, seperatedTop.Ten)
Top.Ten <- Top.Ten[order((Top.Ten$Year)),]
Top.Ten <- Top.Ten[order((Top.Ten$Rank)),]
Top.Ten <- Top.Ten[order(stringr::str_length(Top.Ten$Rank)),]
#Top.Ten

Top.Ten[Top.Ten$Rank==1,]
#MyKuali ranked #1 2014 to 2015 

```


5) (3 pts) Which brand has the highest average star rating? If there is a tie, report multiple brands.

24 brands tie in at 5 stars (ChoripDong, Daddy, Daifuku...). Refer to the data frame below.

```{r}
AvgStarRatings <- aggregate( ramen$Stars ~ ramen$Brand, ramen, mean )
AvgStarRatings[AvgStarRatings$`ramen$Stars`==5.0,]

#Abbie's method
ramen_avgs <- data.frame(brands = as.character(unique(ramen$Brand)), avg_ratings=NA)
for(i  in ramen_avgs$brand){ramen_avgs[ramen_avgs$brands==i, "avg_ratings"] <- mean(ramen[ramen$Brand==i, "Stars"], na.rm=T)}
ramen_avgs

#Abbie's alternative method
ramen_avgs <- aggregate(Stars ~ Brand, data=ramen, mean)
ramen_avgs

###An unasked question

winning_brands <- ramen_avgs[ramen_avgs$Stars == max(ramen_avgs$Stars, na.rm=F), "Brand"]
length(winning_brands)
#length(ramen[ramen$Brand %in% winning_bran
```

6) (5 pts) Is the way a ramen is packaged related to which country it is from?

There is no correlation between packaging style and a country's level of cultural diversity or development, as observed in the plot below. But, it is obvious that pack, cup, and bowl are widely pertinent in all countries (in descending order of significance). Notably, Canada and Australia use less packs than cups.

```{r}
#names(ramen)
#head(ramen)
#Style indicates packaging
#data.frame(sort(unique(ramen$Style)))
#Bar, bowl, box, can, cup, pack, tray

#is packaging correlated to country of origin?
install.packages("ggpubr")
library("ggpubr")
#ggscatter(ramen, x = "Style", y = "Country", 
          #add = "reg.line", conf.int = TRUE, 
          #cor.coef = TRUE, cor.method = "pearson",
          #xlab = "Country of Origin", ylab = "Packaging Style")

#another way, too

table(ramen$Style, ramen$Country)

```


7) (5 pts) What is the average, maximum, and minimum rating for ramens from the US? What about for
ramens from Singapore? Based on your findings, which country do you expect to land in the Top Ten
more often? Check your prediction.
```{r}
mean(ramen[ramen$Country == "USA", "Stars"],  na.rm = T )
max(ramen[ramen$Country == "USA", "Stars"],  na.rm = T )
min(ramen[ramen$Country == "USA", "Stars"],  na.rm = T )

mean(ramen[ramen$Country == "Singapore", "Stars"],  na.rm = T )
max(ramen[ramen$Country == "Singapore", "Stars"],  na.rm = T )
min(ramen[ramen$Country == "Singapore", "Stars"],  na.rm = T )

###Based on my findings Singapore would appear in Top Ten rankings more often because their ramen's mean is higher than USA's

(ramen[ramen$Country == "Singapore", "Top.Ten"])
(ramen[ramen$Country == "USA", "Top.Ten"])

###My prediction is true. Singapore's ramen appears in the top ten 7 times while USA's appears only once


#Abbie's method -- unfinished
#(ramen[ramen$Country=="Singapore" & ramen$Top.Ten !="","Stars"])
#"Stars" & ramen$Top.Ten !="",
```


8) (5 pts) Which country makes the most ramen? Which country makes the best ramen? How did you
define "best"? Why? Was the country that makes the best ramen also the country that makes the
most ramen?
```{r}
sort(table(ramen$Country), decreasing=TRUE)
#Japan make the most ramen

w <- table(ramen$Country, ramen$Stars==5)
#w
sort(w[,2], decreasing = TRUE)
#Japan has the most 5 start rating
# In conclusion Japan make the most ramen and the best ramen
```


9) (5 pts) How many ramens are considered spicy as part of their variety (hint, don’t forget that "spicy""
and "Spicy"" are both spicy)? Are spicy ramens typically rated higher than non-spicy ramens?
```{r}
library(stringr)
sum(str_count(ramen$Variety, "Spicy"))+
  sum(str_count(ramen$Variety, "spicy"))
#277
```


10) (5 pts) Based on this data set, what effect does saltiness have on ramen ratings?

Higher percentage salt content is significantly correlated with ramen ratings. The plot below illustrates and the correlation test confirms; note the low p-value and >0.5 correlation coefficient.

```{r}
plot(ramen$perc_salt, ramen$Stars)

cor.test(ramen$perc_salt, ramen$Stars)
```


11) (5 pts) Are there certain styles, countries, brands, or flavors of ramen that tend to use more or less
salt? Feel free to be creative in how you parse "variety" into flavors.
```{r}
mean(ramen$perc_salt)
SaltRatingsbyStyle <- aggregate(ramen$perc_salt ~ ramen$Style, ramen, mean )
SaltRatingsbyStyle[SaltRatingsbyStyle$`ramen$perc_salt`>19,]
#The countries that use salt more than the mean 
SaltRatingsbyStyle[SaltRatingsbyStyle$`ramen$perc_salt`<19,]
#The countries that use salt less than the mean 

SaltRatingsbyCountry <- aggregate( ramen$perc_salt ~ ramen$Country, ramen, mean )

SaltRatingsbyCountry[SaltRatingsbyCountry$`ramen$perc_salt`>19,]
#The countries that use salt more than the mean 
SaltRatingsbyCountry[SaltRatingsbyCountry$`ramen$perc_salt`<19,]
#The countries that use salt less than the mean 

SaltRatingsbyBrand <- aggregate( ramen$perc_salt ~ ramen$Brand, ramen, mean )
head(SaltRatingsbyBrand[SaltRatingsbyBrand$`ramen$perc_salt`>19,])
#The Brands that use salt more than the mean 
head(SaltRatingsbyBrand[SaltRatingsbyBrand$`ramen$perc_salt`<19,])
#The Brands that use salt more than the mean 


#unique(ramen$Variety)
#install.packages("ggpubr")
#library(ggpubr)
ramen$flavor<-NA  
#ramen
ramen$Variety <- str_to_lower(ramen$Variety)

ramen$flavor[grep(c("curry"), ramen$Variety)] <- "curry"
ramen$flavor[grep(c("tomato"), ramen$Variety)] <- "tomato"
ramen$flavor[grep(c("sesame"), ramen$Variety)] <- "sesame"
ramen$flavor[grep(c("toona"), ramen$Variety)] <- "toona"
ramen$flavor[grep(c("udon"), ramen$Variety)] <- "udon"
ramen$flavor[grep(c("kimchi"), ramen$Variety)] <- "kimchi"
ramen$flavor[grep(c("beef"), ramen$Variety)] <- "beef"
ramen$flavor[grep(c("chicken"), ramen$Variety)] <- "chicken"
ramen$flavor[grep(c("shoyu"), ramen$Variety)] <- "shoyu"
ramen$flavor[grep(c("miso"), ramen$Variety)] <- "miso"
ramen$flavor[grep(c("laksa"), ramen$Variety)] <- "laksa"
ramen$flavor[grep(c("mushroom"), ramen$Variety)] <- "mushroom"
ramen$flavor[grep(c("tempura"), ramen$Variety)] <- "tempura"
ramen$flavor[grep(c("yakisoba"), ramen$Variety)] <- "yakisoba"
ramen$flavor[grep(c("seafood", "prawn", "prawns", "shrimp", "shrimps"), ramen$Variety)] <- "seafood"
#ramen

SaltRatingsbyFlavor <-aggregate( ramen$perc_salt ~ ramen$flavor, ramen, mean )
SaltRatingsbyFlavor[SaltRatingsbyFlavor$`ramen$perc_salt`>19,]
#The Flavor that use salt more than the mean 
SaltRatingsbyFlavor[SaltRatingsbyFlavor$`ramen$perc_salt`<19,]
#The Flavor that use salt less than the mean 


```


```{r}
mean(ramen$perc_salt)
AvgSaltRatings <- aggregate( ramen$perc_salt ~ ramen$Country, ramen, mean )

#there are certain ramen from certain country, brand that uses more salt than other.


#plot(ramen$perc_salt, ramen$Country)
#plot(data.frame(ramen$Country, ramen$perc_salt))
#plot(data.frame(ramen$Brand, ramen$perc_salt))

```


12) [Open Ended] (10 pts) You have been hired by a new ramen start up, Roamin' Ramen. They want to
better understand the ramen market and ask you to break down ramen into 5 collections of "similar" 
ramens. How do you go about it? Discuss what you find.

Key varieties: curry, tomato, sesame, toona, udon, kimchi, beef, chicken, seafood/prawns/shrimps, tempura, mushroom, shoyu, laksa, yakisoba
Main 5 Collections: Chicken, Beef, Seafood, Curry, Udon


There are 5 main kinds of ramen. The japanese would normally say: Shoyu, Shio, Miso, Curry. But, globally, the distinct kinds are actually chicken, beef, seafood, curry, and udon. The code below parses varieties for distinct flavors and counts them. Refer below for the actual results. 

```{r}
#unique(ramen$Variety)

sort(table(ramen$flavor), decreasing=TRUE)

#1 - chicken, 2 - beef, 3 - seafood, 4 - curry, 5 - udon

```


13) [Open Ended] (10 pts) Roamin' Ramen wants to make the next Top Ten Ramen winner. What
recommendations do you have for them based on this data set? Be sure to use the columns provided,
and generate your own ideas by examining the Variety column. Support your assertions with statistics
(like mean, min, max or others you prefer) and at least one plot.

Recommendations: Pack --> Laksa --> Low sodium; Learn from the Japanese and Singaporean ramen houses, relocate to those countries, and/or partner therein

```{r}
ramen$Top.Ten[ramen$Top.Ten==""] <- NA
Top.Ten <- na.omit(ramen)
addtop <- data.frame(do.call('rbind', strsplit(as.character(Top.Ten$Top.Ten), " #", fixed=TRUE)))
colnames(addtop)<-c("Year", "Rank")
Top.Ten <- cbind(Top.Ten, addtop)
Top.Ten$Rank<-as.double(Top.Ten$Rank)
#sort((Top.Ten$Rank), decreasing=FALSE)
#Top.Ten[order(as.numeric(Top.Ten$Rank), decreasing=FALSE),]
#Top.Ten

tab1 <- table(Top.Ten$Rank, Top.Ten$Style)
tab1 <- rbind(tab1, Total = colSums(tab1))
tab1
#14 of the recently ranked Top Ten used "Pack" styles, while there were no "Cup" styles and only 4 "Tray" styles; note, in general pack, cup, and tray styles are most popular in order of decreasing popularity

tab2 <- table(Top.Ten$Rank, Top.Ten$flavor)
tab2 <- rbind(tab2, Total = colSums(tab2))
tab2
#Yakisoba, Laksa, Curry, and Chicken are all strong flavors aong the Top.Ten; Laksa has 2 placements for the #1 title, so we would recommend Laksa as a selected flavor

tab3 <- table(Top.Ten$Rank, Top.Ten$Country)
tab3 <- rbind(tab3, Total = colSums(tab3))
tab3[11,]
#We would recommend learning from, associating with, or partnering with a Japanese or Singaporean ramen house because ramens from those countries consistently rank in the Top Ten

plot(Top.Ten$Rank, Top.Ten$perc_salt, abline(lm(Top.Ten$perc_salt ~ Top.Ten$Rank)))
cor(Top.Ten$perc_salt, Top.Ten$Rank)
#Finally, saltiness will have a 40% impact on the rating, and there is positive correlation between higher ranks and lower saltiness. So, we would recommend reducing percentage salt to a level of 18% to maximize Top Ten likelihood(based on the line of best fit in the plot below).


```
