---
title: "BSDS 100: Intro to Data Science with R Assignment 8"
author: "Chase Darlington (University of San Francisco)"
date: "11/13/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Directions: For all questions in this assignment, write complete sentences and fully answer any question that is asked, and use R to answer each question. Provide all R code and solutions by knitting your final RStudio file into a single file named [your name] CA8.pdf. Late assignments will automatically have 10 points deducted, and will not be accepted after the answer key is posted (1 week following the due date).

```{r packages, include=FALSE}
#install.packages("XML")
#install.packages("rvest")
#install.packages("tidyverse")
library(RCurl)
library(XML)
library(rvest)
library(tidyverse)
```

1. From <https://r4ds.had.co.nz/strings.html> Create regular expressions to find all words from stringr::words that:

#SETUP
```{r}
words <- stringr::words
str(words)
words <- data.frame(words)
nvowels <- str_count(tolower(words[,1]), "[aeoiu]")
total_letters <- str_count(tolower(words[,1]), "\\w")
starts_with_vowel <- grepl("^[aeiou]", tolower(words[,1]))
ends_with_vowel <- grepl("[aeiou]$", tolower(words[,1]))
ends_with_ed <- grepl("ed$", tolower(words[,1]))
ends_with_eed <- grepl("eed$", tolower(words[,1]))
ends_with_ed_not_eed <- ends_with_ed - ends_with_eed
ends_with_ing <- grepl("ing$", tolower(words[,1]))
ends_with_ise <- grepl("ise$", tolower(words[,1]))
ends_with_ing_or_ise <- ends_with_ing + ends_with_ise
words <- cbind(words,nvowels, total_letters, starts_with_vowel, ends_with_ed_not_eed, ends_with_ing_or_ise)
```
(a) (2 points) Start with a vowel.
```{r}
head(words[words$starts_with_vowel==TRUE,])
tail(words[words$starts_with_vowel==TRUE,])
```
(b) (2 points) That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r}
words[words$nvowels==0,]
#technically y is a vowel in these words, but I believe this answers the question
```
(c) (2 points) End with ed, but not with eed. 
```{r}
words[words$ends_with_ed_not_eed==TRUE,]
```
(d) (2 points) End with -ing or -ise.
```{r}
words[words$ends_with_ing_or_ise==TRUE,]
```
2. (4 points) Write and test a regular expression that will match a phone number in the format 555-555-5555

This variable complexity function is more efficient than a for loop, because it executes only as many commands as necessary to invalidate/validate a US number in the format 555-555-5555 (admittedly, I can think of a few quicker ways, especially if you strip out the "-")
```{r}
US_phone_validation <- function(x){
  if(nchar(mynum)==12){
    hyphen1 <- substr(mynum, start = 4, stop = 4)
    hyphen2 <- substr(mynum, start = 8, stop = 8)
    areacode <- substr(mynum, start = 1, stop = 3)
    numpart1 <- substr(mynum, start = 1, stop = 3)
    numpart2 <-substr(mynum, start = 1, stop = 3)
    if(hyphen1=="-"){
      if(hyphen2=="-"){
        if(is.na(as.integer(areacode))==F){
          if(is.na(as.integer(numpart1))==F){
            if(is.na(as.integer(numpart2))==F){
              print("Thank you for your input. We will reach out to you soon.")
            }else{print("Please edit your entry to match the format: 555-555-5555")}
          }else{print("Please edit your entry to match the format: 555-555-5555")}
        }else{print("Please edit your entry to match the format: 555-555-5555")}
      }else{print("Please edit your entry to match the format: 555-555-5555")}
    }else{print("Please edit your entry to match the format: 555-555-5555")}
  }else{print("Please edit your entry to match the format: 555-555-5555")}
}

mynum <- ("253-948-2507")
US_phone_validation(mynum)

mynum <- ("2539482507")
US_phone_validation(mynum)
```

3. (6 points) Use for() and if() to loop through the numbers 1 to 30 and print ”divisible by 3” for any that are divisible by 3 and print ”not divisible by 3” for any that are not divisible by 3
```{r}
#one way
numbers <- 1:30  
is_divisible_by_3 <- NA
num <- data.frame(cbind(numbers, is_divisible_by_3))
for(i in 1:length(numbers)){
  j <- num$numbers[i]
  if(num$numbers[j]%%3==0){
    num$is_divisible_by_3[i] <- "divisible by 3"
    }
  else{
      num$is_divisible_by_3[i] <- "not divisible by 3"
    }
}
num
```

```{r}
#another way
n <- 1:30
for(i in 1:length(n)){
  if(n[i]%%3==0){
    print(paste(n[i], "is divisible by 3", sep=" "))
    }
  else{
    print(paste(n[i], "is not divisible by 3", sep=" "))
    }
}
```


4. (6 points) Now make a vector of the numbers 1 to 30, and use an ifelse() statement to return a vector where 1 is returned if the original number was divisible by 3 and 0 if it was not
```{r}
n <- 1:30
vector <- ifelse(n%%3==0, 
       1,
       0)
matrix <- as.matrix(ifelse(n%%3==0, 
       1,
       0))
  
combined_vector <- rbind(n, ifelse(n%%3==0, 
       1,
       0))

vector
matrix
combined_vector
```
5. (3 points) Consider the following loop. What is wrong with it? What could be done to fix it?
n <- 0
while(n < 10){ print("hello world") n <- n - 1 }

When properly formatted, the function provided is an infinite loop. The while function prints so long as n is less than 10, and counts n down (n <- n-1) at the same time. Subsequently, n is continuously less than 10, and the function runs an infinite number of times.

```{r}
n <- 0
while(n < 10){
  print("hello world") 
  n = n+1
  }
```
To resolve the function, simply count n up (n <- n+1), and it will no longer run an infinite number of times.

6. (3 points) Write and test a switch function that returns ”woof” when passed ”dog”, ”meow” when passed ”cat”, and ”???” when passed ”fox”
```{r}
n <- "dog"
switch(n, "dog"="woof", "cat"="meow", "fox"="???")

n <- "cat"
switch(n, "dog"="woof", "cat"="meow", "fox"="???")

n <- "fox"
switch(n, "dog"="woof", "cat"="meow", "fox"="???")
#What does the fox say?
```
