---
title: "Chase Darlington_CA3"
author: "Chase Darlington"
date: "September 4, 2018"
output: pdf_document
---
#Load the data

```{r}
data(iris)
```

#Look at the data
```{r}
iris
```

#Read a description of the data

```{r}
help(iris)
```

#Display the column and row names of the data

```{r}
colnames(iris)
rownames(iris)
```

#Look at the 10th row

```{r}
iris[10, ]
```

#Look at the 3rd column


```{r}
iris[, 3]
```

#Alternatively, just look at the variable "Petal.Length"

```{r}
iris$Petal.Length
```

#Store columns 1-2 and rows 10-20 for later use

```{r}
subset.data <- iris[10:20, 1:2]
```

#Store the species names for later use

```{r}
Species.names <- iris$Species
```

# (a) (5 points) What are the different variables in this dataset? What are the measurements of each of these variables for the 10th sample?

```{r}
help(iris)
```


# 5 variables (columns) named Sepal.Length, Sepal.Width, Petal.Length, Petal.Width, and Species.

```{r}
iris[10,]
```

     
#   (b) (5 points) How many samples are in this dataset? What are the different species of flower that have been measured?

```{r}
help(iris)
```


# OR

```{r}
rownames(iris)
```


#sample-size: 150 cases

```{r}
iris[,5]
```

# OR

```{r}
print(Species.names)
```


# species: setosa, versicolor, virginica


#calculate the 5-number summary of the {\it iris} dataset

```{r}
five.num.summary <- summary(iris)
print(summary(iris))
```

#calculate the standard deviation across the first four columns of the dataset

```{r}
st.devs.cols <- apply(iris[,1:4], 2, sd)
print(st.devs.cols)
```

#calculate the standard deviation across the samples of the dataset

```{r}
st.devs.rows <- apply(iris[, 1:4], 1, sd)
print(st.devs.rows)
boxplot(iris[, 1:4], main = "Boxplot of Iris Variables")
```


# (a) (5 points) How many of each species are there in the dataset?

```{r}
five.num.summary <- summary(iris)
print(summary(iris))
```


# 50 of each specie
#   (b) (5 points) Which two variables have a median that is smaller than their corresponding mean

```{r}
five.num.summary <- summary(iris)
print(summary(iris))
```

# Sepal.Length and Sepal.Width
#   (c) (5 points) What is the standard deviation of the sepal length measurements?

```{r}
st.devs.cols <- apply(iris[,1:4], 2, sd)
print(st.devs.cols)
```

# sd of Sepal.Length=0.8280661

#   (d) (5 points) Calculate the 5-number summary and standard deviations of the subset you extracted earlier (subset.data). What is the standard deviation of the sepal length measurements for this subset?

```{r}
summary(subset.data)
apply(subset.data,2, sd)
```


# sd of Sepal Length in subset=0.4665151


